---
title: "Cheat Sheet is GOOD "
author: "Renjie Wei"
date: "10/29/2020"
output: 
  html_document:
    toc: true
    toc_float : true
    code_folding: hide
---

# Discrete Distribution

## Binomial
$$X \sim Bin(n,p)$$

```{r binomial}

```
$$P({X}=k)\ = \binom{n}{k} p^k(1-p)^{n-k}$$
$$P(X \leq k)=\sum_{i=0}^{k}\binom{n}{i}p^i(1-p)^{n-i}$$

## Poisson

$$X \sim Poi(\lambda)$$

```{r poisson}

```
$$P(X = k) = \frac{\lambda^{k}e^{-\lambda}}{k!}$$
$$P(X \leq k) = \sum_{i=0}^{k}\frac{\lambda^{i}e^{-\lambda}}{i!}$$

# Continuos Distribution

## Uniform Distribution

$$X\sim Unif(a,b)$$

The pdf of uniform distribution is:

$$f(x) = \frac{1}{b-a}$$
$$P(X\leq k) = P(X<k)= \int_{a}^{k}f(x)dx = \int_{a}^{k}\frac{1}{b-a}dx,~(a\leq k \leq b) $$
```{r unif}

```


## Normal Distribution

$$X \sim(\mu,\sigma^2)$$
The pdf of normal distribution is:
$$f(x) = \frac{1}{\sigma\sqrt{2\pi}}e^{\frac{-(x-\mu)^2}{2\sigma^2}}$$
The standard normal distribution:
$$Z=\frac{X-\mu}{\sigma}\sim N(0,1)$$
$$\phi(z) = \frac{1}{\sqrt{2\pi}}e^{\frac{-z^2}{2}}$$

$$P(X\leq k) = P(Z \leq \frac{k-\mu}{\sigma}) = P(Z \leq Z_{score})$$
$$Z_\alpha = P(Z <Z_\alpha) = \alpha$$
```{r norm}

```

## Normal Approximation to Binomial and Poisson

Rule of Thumb:
$$np \geq 10, ~n(1-p) \geq 10$$
$$X\sim N(np,np(1-p))~or~ X\sim(\lambda,\lambda)$$



Continuity Correction:
$$to~ be~ continue$$


## Student's t-Distribution

$$T\sim t(v)$$


# One Sample Mean

## Interval Estimation

Construct the 95% CI for the true mean:
$$
\begin{split}
n=10\\
\sigma =15\\
\bar X = 175\\
\alpha = 5\%\\
95\% CI: \bar X \pm Z_{1-\alpha/2} \cdot \sigma/ \sqrt n\\ 
= 175 \pm Z_{0.975}\cdot15/\sqrt{10} \\
= (175-1.96\cdot 15/\sqrt{10},175+1.96\cdot 15/\sqrt{10})\\
=(165.70,184.30)
\end{split}
$$

## Test for the Mean of a Normal Distribution with Known Variance


$$X\sim N(\mu,\sigma^2)$$

### One-sided Test

* left tailed

$$
\begin{split}
H_0 : \mu=\mu_0\\
H_1 : \mu<\mu_0\\
\bar X=\\
n = \\
\sigma = \\
\alpha = \\
Z_{stat}= \frac{\bar X - \mu_0}{\sigma/\sqrt{n}}\\
Z_{critical}=Z_{\alpha} \\
Z_{stat} \sim N(0,1) \ under \ the \ null
\end{split}
$$

**Decision Rule:**

reject $H_0$ if $Z_{stat} < Z_{\alpha}$

Fail to reject $H_0$ otherwise.

P-value = $Pr(Z< Z_{stat} | H_0)$


At 5% significance level, $Z<Z_{critical}$, $Pval<$, we reject the null and conclude that there is a significant difference in ...


* Right tailed

$$
\begin{split}
H_0 : \mu=\mu_0\\
H_1 : \mu>\mu_0\\
\bar X=\\
n = \\
\sigma = \\
\alpha = \\
Z_{stat}= \frac{\bar X - \mu_0}{\sigma/\sqrt{n}}\\
Z_{critical}=Z_{1-\alpha} \\
Z_{stat} \sim N(0,1) \ under \ the \ null
\end{split}
$$

**Decision Rule:**

reject $H_0$ if $Z_{stat} > Z_{1-\alpha}$

Fail to reject $H_0$ otherwise.

P-value = $Pr(Z > Z_{stat} | H_0)$


At 5% significance level, $Z_{stat}>Z_{critical}$, $Pval<$, we reject the null and conclude that there is a significant difference in ...


### Two-sided Test

$$
\begin{split}
H_0 : \mu=\mu_0\\
H_1 : \mu \neq \mu_0\\
\bar X=\\
n = \\
\sigma = \\
\alpha = \\
Z_{stat}= \frac{\bar X - \mu_0}{\sigma/\sqrt{n}}\\
Z_{critical}=Z_{1-\alpha/2} \\
Z_{stat} \sim N(0,1) \ under \ the \ null
\end{split}
$$

**Decision Rule:**

reject $H_0$ if $|Z_{stat}| > Z_{1-\alpha/2}$

Fail to reject $H_0$ otherwise.

P-value = $2\times Pr(Z > Z_{stat} | H_0)$, if $Z_{stat} \geq 0$, or

P-value = $2\times Pr(Z < Z_{stat} | H_0)$, if $Z_{stat} \leq 0$


At 5% significance level, $|Z_{stat}|>Z_{critical}$, $Pval<$, we reject the null and conclude that there is a significant difference in ...


## Test for the Mean of a Normal Distribution with Unknown Variance

### One-sided Test

* left tailed

$$
\begin{split}
H_0 : \mu=\mu_0\\
H_1 : \mu<\mu_0\\
\bar X=\\
n = \\
s = \\
\alpha = \\
t_{stat}= \frac{\bar X - \mu_0}{s/\sqrt{n}}\\
t_{critical}=t_{n-1,\alpha} \\
t_{stat} \sim T(n-1) \ under \ the \ null
\end{split}
$$

**Decision Rule:**

reject $H_0$ if $t_{stat} < t_{\alpha}$

Fail to reject $H_0$ otherwise.

P-value = $Pr(t_{n-1} < t_{stat} | H_0)$


At 5% significance level, $t_{stat}<t_{critical}$, $Pval<$, we reject the null and conclude that there is a significant difference in ...


* Right tailed

$$
\begin{split}
H_0 : \mu=\mu_0\\
H_1 : \mu>\mu_0\\
\bar X=\\
n = \\
s = \\
\alpha = \\
t_{stat}= \frac{\bar X - \mu_0}{s/\sqrt{n}}\\
t_{critical}=t_{n-1,1-\alpha} \\
t_{stat} \sim T(n-1) \ under \ the \ null
\end{split}
$$

**Decision Rule:**

reject $H_0$ if $t_{stat} > t_{1-\alpha}$

Fail to reject $H_0$ otherwise.

P-value = $Pr(t_{n-1} > t_{stat} | H_0)$


At 5% significance level, $t_{stat}>t_{critical}$, $Pval<$, we reject the null and conclude that there is a significant difference in ...


### Two-sided Test

$$
\begin{split}
H_0 : \mu=\mu_0\\
H_1 : \mu \neq \mu_0\\
\bar X=\\
n = \\
s = \\
\alpha = \\
t_{stat}= \frac{\bar X - \mu_0}{s/\sqrt{n}}\\
t_{critical}=t_{n-1,1-\alpha/2} \\
t_{stat} \sim T(n-1) \ under \ the \ null
\end{split}
$$

**Decision Rule:**

reject $H_0$ if $|t_{stat}| > t_{n-1,1-\alpha/2}$

Fail to reject $H_0$ otherwise.

P-value = $2\times Pr(t_{n-1} > t_{stat} | H_0)$, if $t_{stat} \geq 0$, or

P-value = $2\times Pr(t_{n-1} < t_{stat} | H_0)$, if $t_{stat} \leq 0$


At 5% significance level, $|t_{stat}|>t_{critical}$, $Pval<$, we reject the null and conclude that there is a significant difference in ...

# Two Sample Mean

## Paired t-test

This is a sample from hw3.

$$
\begin{split}
H_0 = \mu_{post} - \mu_{pre} = \Delta = 0 \\
H_1 = \mu_{post} - \mu_{pre} = \Delta \neq 0 \\
d_1 = -8.57 \\
S_{d1} = 17.17 \\
n = 36\\
t_{stat}= \frac{\bar{d_1}}{S_{d1}/\sqrt{n}}=\frac{-8.58}{17.17/\sqrt{36}} = -2.99 \\
t_{critical}=t_{n-1, 1-\alpha/2}= t_{35, 0.975}=2.03 \\
t_{stat} \sim t_{35} \ under \ the \ null
\end{split}
$$

**Decision Rule:**

reject $H_0$ if $|t_{stat}| > t_{35, 0.975}$

Fail to reject $H_0$ otherwise.

Pvalue = $2* Pr(t_{35} < t_{stat} | H_0) = 2* Pr(t_{35} < -2.99 | H_0) = 0.005$


At 5% sifnificance level, $|t_{stat}|>t_{critical}$, $Pval<0.05$, we reject the null and conclude that there is a significant difference in ...


$$CI = \bar d \pm t_{n-1, 1-\alpha/2} \times s_d/ \sqrt n$$ 

## Test for Equality of Variance

$$
\begin{split}
H_0 = \sigma^2_{d1} = \sigma^2_{d0} \\
H_1 = \sigma^2_{d1} \neq \sigma^2_{d0} \\
F_{stat}= \frac {s^2_{d1}}{s^2_{d0}} \sim F_{35,35} \ Under \ the \ Null \\
F_{stat}= \frac {17.17^2}{14.81^2} = 1.34 \\
F_{critical} = F_{35, 35, 0.975} = 1.96
\end{split}
$$

**Decision Rule:**

reject $H_0$ if $F_{stat} > F_{35, 35, 0.975}$

Fail to reject $H_0$ otherwise

Pvalue = $2* Pr(F_{n_1-1,n_2-1} < F_{stat} | H_0),if~ F<1$, or $2* Pr(F_{n_1-1,n_2-1} > F_{stat} | H_0),if~ F\geq 1$

At 5% significance level, $F_{stat}< F_{critical}$, we fail to reject the null and conclude that there is not enough evidence to support a difference in the population variances.


## Two-sample independent test, equal variance

$$
\begin{split}
H_0 = \mu_{d1} - \mu_{d0} = 0 \\
H_1 = \mu_{d1} - \mu_{d0} \neq 0
\end{split}
$$


$$
\begin{split}
\bar{X_1} - \bar{X_2} \sim N(\mu_1-\mu_2,\frac{\sigma^2}{n_1}+\frac{\sigma^2}{n_2}) ~if~ the~ population~ variance~ is~ given\\
otherwise,\\
t_{stat} = \frac {\bar{X_{d1}} - \bar{X_{d0}}} {s\sqrt {1/n_1 + 1/n_0}} \sim t_{n_1+n_0-2, 1-\alpha/2} \\
where \ s= \sqrt{\frac{(n_1-1)s_1^2+(n_0-1)s_0^2}{n_1+n_2-2}} \\
s = \sqrt{\frac{(35)(219.34)+(35)(294.81)}{70}} = 16.03 \\
t_{stat} = \frac {-8.58 + 3.33}{(16.03)\sqrt {1/36 + 1/36}} = -1.39 \\
t_{critical} = t_{n_1+n_0-2, 1-\alpha/2} = t_{70, 0.975} = 1.99
\end{split}
$$
**Decision Rule:**

reject $H_0$ if $|t_{stat}| > t_{critical}$

Fail to reject $H_0$ otherwise.

Pvalue = $2* Pr(t_{70} < t_{stat} | H_0) =2* Pr(t_{70} < -1.39 | H_0)$ = 0.17

At 5% significance level, $|t_{stat}|<t_{critical}$, $Pval>0.05$, we fail to reject the null and conclude that there is not enough evidence to support a difference in the mean SBP between the intervention group and control group.  

This conclusion is supported by the 95% confidence interval since 0 is contained within the interval:  

$$ CI = \bar{X_{d1}} - \bar{X_{d0}} \pm t_{n_1+n_0-2, 1-\alpha/2} \times s \sqrt {1/n_1 + 1/n_0} = (-8.58 + 3.33) \pm 1.99 \times 16.03 \sqrt{1/36 + 1/36} = (-12.77, 2.27)$$

## Two-sample independent test, unequal variance

$$
\begin{split}
\bar{X_1} - \bar{X_2} \sim N(\mu_1-\mu_2,\frac{\sigma_1^2}{n_1}+\frac{\sigma_2^2}{n_2}) ~if~ the~ population~ variance~ is~ given\\
otherwise,\\
t_{stat} = \frac{\bar{X_1} - \bar{X_2}}{\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}}}~\sim t_{d''}\\
d'' =[d'],~where,\\
d' = \frac{(\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2})^2}{\frac{s_1^2}{n_1}^2/(n_1-1)+\frac{s_2^2}{n_2}^2/(n_2-1)}
\end{split}
$$

# ANOVA


$$
\begin{split}
H_0: \mu_1=\mu_2=...=\mu _k\\
H_1: at~least~two~means~are~not~equal\\
Between~SS = \sum_{i=1}^k\sum_{j=1}^{n_i}(\bar{y_i} - \bar{\bar{y}})^2\\
Within~SS = \sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}-\bar{y_i})^2\\
Between~Mean~Square = \frac{\sum_{i=1}^k\sum_{j=1}^{n_i}(\bar{y_i} - \bar{\bar{y}})^2}{k-1}=\sum_{i=1}^kn_i\bar{y_i}^2-\frac{y_{..}^2}{n}\\
Within~Mean~Square = \frac{\sum_{i=1}^k\sum_{j=1}^{n_i}(y_{ij}-\bar{y_i})^2}{n-k}=\sum_{i=1}^k(n_i-1)s_i^2\\
F_{stat} = \frac{Between~Mean~Square}{Within~Mean~Square} \sim F(k-1,n-k)
\end{split}
$$


**Decision Rule:**

Reject $H_0$ if $F>F_{k-1,n-k,1-\alpha}$

Fail to reject $H_0$ otherwise.

Pvalue = $Pr(F_{k-1,n-k}>F_{stat})$


# One Sample Prop

$$\bar X \sim N(p,\frac{p(1-p)}{n})$$

$$
\begin{split}
H_0:p=p_0\\
H_1: p \neq p_0\\
\hat p = \frac{\sum_{i=1}^{n}X_i}{n}=\frac{X}{n}\\
Z_{stat} = \frac{\hat p -p_0}{\sqrt{p_0(1-p_0)/n}}
\end{split}
$$

**Decision Rule:**

Reject $H_0$


# Two Sample Prop




# Chi_Sq

# Fisher's Exact and McNemar
